raster_brick_stuff

### 2018-10-03 src/biological_data_prep.Rmd
warm/cold year intro
more raster loop mess...
depth layer codes sorted!

# Column removal
So there are many columns that you don't need.. you want to keep
id
decimalLatitude
decimalLongitude
datecollected
institutioncode
individualcount
resname
originalscientificname
collectioncode (this contains some info that might help with finding out gear used in the survey)

```{r select colums needed}
tidydata <- subset(tidydata, select = c(id, decimalLatitude, decimalLongitude, datecollected, institutioncode, individualcount, resname, originalscientificname, collectioncode))
head(tidydata)
```

@@

>>>>

# Column removal
So there are many columns that you don't need.. you want to keep
id
decimalLatitude
decimalLongitude
datecollected
institutioncode
individualcount
resname
originalscientificname
collectioncode (this contains some info that might help with finding out gear used in the survey)
depth

```{r select colums needed}
tidydata <- subset(tidydata, select = c(id, decimalLatitude, decimalLongitude, datecollected, institutioncode, individualcount, depth, resname, originalscientificname, collectioncode))
head(tidydata)
```

@@

# depth bins
the depth layers in the netCDF are literal slices. I want to attribute the observations with depth data to the nearest depth layer. I have created a .csv which lists all the depth layers, and then created depth bins with min and max values.

load the depth bin csv
```{r}
depthbin <- read.csv("../data/env/depth_bin.csv", header = TRUE) 
head(depthbin)

so what you want to do is look at the observation, see if it has a depth value (data_aea$depth) and if so, find the value that lies between depthbin$bin_min and depthbin$bin_max, then extract depthbin$layer_name and add to data_aea$depth_layer, and then extract depthbin$layer_no and add to data_aea$depthlayer_no

first make sure that the data_aea$depth without values == NA (currently they are marked xx)
```{r}
data_aea$depth[data_aea$depth == "xx"] <- NA #make all xx values NA
```

the depth data in data_aea is a factor. This needs to be numeric
```{r}
data_aea$depth <- as.numeric(as.character(data_aea$depth))
str(data_aea)
```

Now run some fancy if loop (courtesy of [TinglTanglBob] (https://stackoverflow.com/questions/52626995/problem-with-loop-to-add-values-from-one-dataframe-and-to-another-based-on-condi/52627728#52627728))
```{r}
check_depth <- function(d_temp)
{
  #print(d_temp)
  if(is.na(d_temp)) return(NA) # if d_temp is na just return NA 
  layer_name_temp <- depthbin$layer_name[which(depthbin$bin_min <= d_temp & depthbin$bin_max >= d_temp)]
  if(length(layer_name_temp) > 1) layer_name_temp <- layer_name_temp[1] # in case there are more hits, the first one is taken
  return(layer_name_temp)
}

data_aea$depth_layer <- sapply(data_aea$depth, check_depth)
```

ok now dow the same thing to get your layer_no
```{r}
check_depth2 <- function(d_temp)
{
  #print(d_temp)
  if(is.na(d_temp)) return(NA) # if d_temp is na just return NA 
  layer_no_temp <- depthbin$layer_no[which(depthbin$bin_min <= d_temp & depthbin$bin_max >= d_temp)]
  if(length(layer_no_temp) > 1) layer_no_temp <- layer_no_temp[1] # in case there are more hits, the first one is taken
  return(layer_no_temp)
}

data_aea$depthlayerno <- sapply(data_aea$depth, check_depth2)
```

# warm/cold year

The warm/cold year index will be taken from a [DFO CSAS report](http://www.dfo-mpo.gc.ca/csas-sccs/Publications/SAR-AS/2018/2018_039-eng.pdf)

### 2018-10-03 src/environmental_data_preperation.Rmd
depth layer codes sorted!

read in a biomer netcdf
```{r}
library("ncdf4")
biobrick <- brick("../data/env/CMEMS_GLOBAL_REANALYSIS_BIO_001_018_chl_monthly_mean_1998-01.ncwgs.nc_lonlat.ncdepthtrim.nc")
biobrick
```

Project the biomer file to match the glorys layer
```{r}
bio_aea <- projectRaster(biobrick, aea) #note this only projects the top layer
bio_aea
writeRaster(bio_aea,'../output/env/bio_aea.tif',overwrite=TRUE, options=c('TFW=YES')) #so you can access it again
```

>>>>

read in a biomer netcdf
note this only reads in the top layer which is what i need right now. To get the other layers, add ", lvar = 4" to brick (see code)
```{r}
library("ncdf4")
biobrick <- brick("../data/env/CMEMS_GLOBAL_REANALYSIS_BIO_001_018_chl_monthly_mean_1998-01.ncwgs.nc_lonlat.ncdepthtrim.nc") # lvar = 4 
biobrick
```

Project the biomer file to match the glorys layer
```{r}
bio_aea <- projectRaster(biobrick, aea) 
bio_aea
writeRaster(bio_aea,'../output/env/bio_aea.tif',overwrite=TRUE, options=c('TFW=YES')) #so you can access it again
```

### 2018-10-03 src/help_extract_from_multiple_netCDFs.Rmd
depth layer codes sorted!

Small function 1) Reads filename & explicitly opens it as a NetCDF file.
```{r}
open_netcdf_as_rasterbrick <- function( ncdf_filename_input ) {
  ncdf_filename_input %>%
  file.path( weatherraster_path , . ) %>%
  nc_open( . )  %>%  
  ncvar_get( . ) %>%   
  brick( . ) %>% # original code: brick( . ,ymn = -0, ymx = 360, xmn = -90, xmx = 90, crs = "the string for your CRS" ) %>%
  return( . )
}
```

@@

# Small function 3) Extract data from a raster brick.
```{r}
generate_data_from_rasterandshape <- function( brick_input ) {
	brick_input %>%
	# Take means according to the countryshape.
	# Make sure df = TRUE , so that output is a dataframe.
    raster::extract( . , countryshape ,
    				f=TRUE, fun = mean, na.rm = TRUE  )
    return( . )
}
```

# Small function 4) Grab 4-digit year from input filename.
grab_year_from_inputfile <- function( ncdf_filename_input ) {

  ncdf_filename_input %>%
  regexpr("[0-9]+", . ) %>%  # Match 4-digit year.
  regmatches( ncdf_filename_input , . ) %>%  # Get matched REGEX from input string.
  return( . )
}

>>>>

Small function 1) Reads filename & explicitly opens it as a NetCDF file.
```{r}
open_netcdf_as_rasterbrick <- function( ncdf_filename_input ) {
  ncdf_filename_input %>%
  file.path( weatherraster_path , . ) %>%
  nc_open( . )  %>%  
  ncvar_get( . ) %>%   
  brick( . , lvar = 4) %>% # lvar = 4 to read in all depth levels (nlayers)
  return( . )
}
```

@@

# Small function 3) Extract data from a raster brick.
```{r}
generate_data_from_rasterandshape <- function( brick_input ) {
	# this whole section needs to be changed 
  brick_input %>%
	# Take means according to the countryshape.
	# Make sure df = TRUE , so that output is a dataframe.
    raster::extract( . , countryshape ,
    				f=TRUE, fun = mean, na.rm = TRUE  )
    return( . )
}
```

# Small function 4) Grab 4-digit year from input filename.
```{r}
grab_year_from_inputfile <- function( ncdf_filename_input ) {

  ncdf_filename_input %>%
  regexpr("[0-9]+", . ) %>%  # Match 4-digit year.
  regmatches( ncdf_filename_input , . ) %>%  # Get matched REGEX from input string.
  return( . )
}
```

### 2018-10-03 src/raster_brick_stuff.Rmd
depth layer codes sorted!

A note to anyone who might happen to stumble across this... I am a beginner in R and have had no exposure to similar languages. The code herein is unlikely to be elegant and there is likely more efficient ways of running the code.

@@

>>>>

A note to anyone who might happen to stumble across this... I am a beginner in R and have had no exposure to similar languages. I don't know what I'm doing. The code herein is unlikely to be elegant and there area probably more efficient ways of running the code.

@@

```{r}
brickttest <- nc_open("../data/env/netcdftest/1993_01_salinity.nc")
testbrick <- brick( "../data/env/netcdftest/1993_01_salinity.nc", lvar = 4)

```
